### 2019-11-04 submission/manuscript.tex
Incorporating Pat's edits for the response to reviewers

\section{\texorpdfstring{The proton pump inhibitor omeprazole does not
promote \emph{Clostridium difficile} colonization in a murine
model}{The proton pump inhibitor omeprazole does not promote Clostridium difficile colonization in a murine model}}\label{the-proton-pump-inhibitor-omeprazole-does-not-promote-clostridium-difficile-colonization-in-a-murine-model}

@@

Proton pump inhibitor (PPI) use has been associated with microbiota
alterations and susceptibility to \emph{Clostridium}
(\emph{Clostridioides}) \emph{difficile} infections (CDIs) in humans. We
assessed how PPI treatment alters the fecal microbiota and whether
treatment promotes CDIs in a mouse model. Mice receiving a PPI treatment
were gavaged with 40 mg/kg of omeprazole during a 7-day pretreatment
phase, the day of \emph{C. difficile} challenge, and the following 9
days. We found that mice treated with omeprazole were not colonized by
\emph{C. difficile}. When omeprazole treatment was combined with a
single clindamycin treatment, one cage of mice remained resistant to
\emph{C. difficile} colonization, while the other cage was colonized.
Treating mice with only clindamycin followed by challenge resulted in
\emph{C. difficile} colonization. 16S rRNA gene sequencing analysis
revealed that omeprazole had minimal impact on the structure of the
murine microbiota throughout the 16 days of omeprazole exposure. These
results suggest omeprazole treatment alone is not sufficient to disrupt
microbiota resistance to \emph{C. difficile} infection in mice that are
normally resistant in the absence of antibiotic treatment.

@@

Antibiotics are the primary risk factor for \emph{Clostridium}
(\emph{Clostridioides}) \emph{difficile} infections (CDIs), but other
factors may also increase a person's risk. In epidemiological studies,
proton pump inhibitor (PPI) use has been associated with CDI incidence
and recurrence. PPIs have also been associated with alterations in the
human intestinal microbiota in observational and interventional studies.
We evaluated the effects of the PPI omeprazole on the structure of the
murine intestinal microbiota and its ability to disrupt colonization
resistance to \emph{C. difficile}. We found omeprazole treatment had
minimal impact on the murine fecal microbiota and did not promote
\emph{C. difficile} colonization. Further studies are needed to
determine whether other factors contribute to the association between
PPIs and CDIs seen in humans or whether aspects of murine physiology may
limit its utility to test these types of hypotheses.

@@

Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridium}
(\emph{Clostridioides}) \emph{difficile} infections (CDIs) (1). It is
less clear whether other human medications that impact the microbiota
also influence \emph{C. difficile} colonization resistance. Multiple
epidemiological studies have suggested an association between proton
pump inhibitor (PPI) use and incidence or recurrence of CDIs (2--5).
There have also been a number of large cohort studies and interventional
clinical trials that demonstrated specific alterations in the intestinal
microbiome were associated with PPI use (4, 6). PPI-associated
microbiota changes have been attributed to the ability of PPIs to
increase stomach acid pH which may promote the survival of oral and
pathogenic bacteria (4, 6). In human fecal samples, PPI use results in
increases in \emph{Enterococcaceae}, \emph{Lactobacillaceae},
\emph{Micrococcaceae}, \emph{Staphylococcaceae} and
\emph{Streptococcaceae} and decreases in \emph{Ruminococcaceae} (6--9).
Several of these taxa have also been associated with \emph{C. difficile}
colonization in humans (10).\\
Unfortunately, the studies suggesting a link between PPIs and \emph{C.
difficile} were retrospective and did not evaluate changes in the
microbiome (2, 3, 5). Thus, it is unclear whether the gastrointestinal
microbiome changes associated with PPI use explain the association
between PPIs and CDIs. Additionally, epidemiological studies have a
limited capacity to address potential confounders and comorbidities in
patients that were on PPIs and developed CDIs or recurrent CDIs (2, 5).
Here, we evaluated the impact of daily PPI treatment with omeprazole on
the murine microbiome and susceptibility to \emph{C. difficile}
colonization in relation to clindamycin, an antibiotic that perturbs the
microbiome enough to allow \emph{C. difficile} to colonize but is mild
enough that \emph{C. difficile} is cleared within 10 days (11).

>>>>

\section{\texorpdfstring{The proton pump inhibitor omeprazole does not
promote \emph{Clostridioides difficile} colonization in a murine
model}{The proton pump inhibitor omeprazole does not promote Clostridioides difficile colonization in a murine model}}\label{the-proton-pump-inhibitor-omeprazole-does-not-promote-clostridioides-difficile-colonization-in-a-murine-model}

@@

Proton pump inhibitor (PPI) use has been associated with microbiota
alterations and susceptibility to \emph{Clostridioides difficile}
infections (CDIs) in humans. We assessed how PPI treatment alters the
fecal microbiota and whether treatment promotes CDIs in a mouse model.
Mice receiving a PPI treatment were gavaged with 40 mg/kg of omeprazole
during a 7-day pretreatment phase, the day of \emph{C. difficile}
challenge, and the following 9 days. We found that mice treated with
omeprazole were not colonized by \emph{C. difficile}. When omeprazole
treatment was combined with a single clindamycin treatment, one cage of
mice remained resistant to \emph{C. difficile} colonization, while the
other cage was colonized. Treating mice with only clindamycin followed
by challenge resulted in \emph{C. difficile} colonization. 16S rRNA gene
sequencing analysis revealed that omeprazole had minimal impact on the
structure of the murine microbiota throughout the 16 days of omeprazole
exposure. These results suggest omeprazole treatment alone is not
sufficient to disrupt microbiota resistance to \emph{C. difficile}
infection in mice that are normally resistant in the absence of
antibiotic treatment.

@@

Antibiotics are the primary risk factor for \emph{Clostridioides
difficile} infections (CDIs), but other factors may also increase a
person's risk. In epidemiological studies, proton pump inhibitor (PPI)
use has been associated with CDI incidence and recurrence. PPIs have
also been associated with alterations in the human intestinal microbiota
in observational and interventional studies. We evaluated the effects of
the PPI omeprazole on the structure of the murine intestinal microbiota
and its ability to disrupt colonization resistance to \emph{C.
difficile}. We found omeprazole treatment had minimal impact on the
murine fecal microbiota and did not promote \emph{C. difficile}
colonization. Further studies are needed to determine whether other
factors contribute to the association between PPIs and CDIs seen in
humans or whether aspects of murine physiology may limit its utility to
test these types of hypotheses.

@@

Antibiotics have a large impact on the intestinal microbiome and are a
primary risk factor for developing \emph{Clostridioides difficile}
infections (CDIs) (1). It is less clear whether other human medications
that impact the microbiota also influence \emph{C. difficile}
colonization resistance. Multiple epidemiological studies have suggested
an association between proton pump inhibitor (PPI) use and incidence or
recurrence of CDIs (2--5). There have also been a number of large cohort
studies and interventional clinical trials that demonstrated specific
alterations in the intestinal microbiome were associated with PPI use
(4, 6). PPI-associated microbiota changes have been attributed to the
ability of PPIs to increase stomach acid pH which may promote the
survival of oral and pathogenic bacteria (4, 6). In human fecal samples,
PPI use results in increases in \emph{Enterococcaceae},
\emph{Lactobacillaceae}, \emph{Micrococcaceae}, \emph{Staphylococcaceae}
and \emph{Streptococcaceae} and decreases in \emph{Ruminococcaceae}
(6--9). Several of these taxa have also been associated with \emph{C.
difficile} colonization in humans (10).\\
Unfortunately, the studies suggesting a link between PPIs and \emph{C.
difficile} were retrospective and did not evaluate changes in the
microbiome (2, 3, 5). Thus, it is unclear whether the gastrointestinal
microbiome changes associated with PPI use explain the association
between PPIs and CDIs. Additionally, epidemiological studies have a
limited capacity to address potential confounders and comorbidities in
patients that were on PPIs and developed CDIs or recurrent CDIs (2, 5).
Here, we evaluated the impact of daily PPI treatment with omeprazole on
the murine microbiome and susceptibility to \emph{C. difficile}
colonization in relation to clindamycin, an antibiotic that perturbs the
microbiome enough to allow \emph{C. difficile} to colonize but is mild
enough that \emph{C. difficile} is cleared within 10 days (11).
